TEMPERATURES AND METALLICITIES OF M DWARFS IN THE \apogee\ SURVEY

### 2019-07-31 paper/draft.tex
switch order
abstract + title

\title{TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY}

@@

M dwarfs have enourmous potential for understanding structure and
formation on both Galactic and exoplanetary scales
with their properties and compositions.
However, current atmospheric models
have limited abilities at reproducing spectral features in stars of the 
coolest temperature regimes ($T_{\rm eff} < 4200\,$K) and fully
exploiting the information content and statistical extent of
current and upcoming large-scale spectroscopic surveys.
Here we present a catalog of spectroscopic temperatures, metallicities,
spectral types for 5,875 M dwarfs in the \apogee\ survey using
\thecannon: a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs with optically derived labels spanning 
$2860 < \teff < 4130$\,K callibrated with bolometric temperatures, 
$-0.5 < \feh < 0.5$\,dex callibrated with FGK wide binary metallicities, 
we train a two-parameter model with predictive accuracy (in cross-validation) 
to 77\,K and 0.09\,dex respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known SDSS optical spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
\color{gcolor}{From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
apply our models and compare our \cannon\ derived $\teff$ against color--temperature
relations and \cannon\ derived $\feh$ against isochrone models in
\gaia\ color-magnitude space.
Finally we use our results to
estimate stellar radii from absolute magnitudes and \cannon\ temperatures [AND WE FIND WHAT??].}\color{black}
\end{abstract}

>>>>

\title{TEMPERATURES AND METALLICITIES OF M DWARFS IN THE \apogee\ SURVEY}

@@

M dwarfs have enourmous potential for understanding structure and
formation on both Galactic and exoplanetary scales
with their properties and compositions.
However, current atmospheric models
have limited abilities at reproducing spectral features in stars of the 
coolest temperature regimes ($T_{\rm eff} < 4200\,$K) and fully
exploiting the information content and statistical extent of
current and upcoming large-scale spectroscopic surveys.
Here we present a catalog of spectroscopic temperatures, metallicities,
spectral types for 5,875 M dwarfs in the \apogee\ survey using
\thecannon: a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs with optically derived labels spanning 
$2860 < \teff < 4130$\,K callibrated with bolometric temperatures, 
$-0.5 < \feh < 0.5$\,dex callibrated with FGK binary metallicities, 
we train a two-parameter model with predictive accuracy (in cross-validation) 
to 77\,K and 0.09\,dex respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known SDSS optical spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
apply our models to a selection of 5,875 M dwarfs. 
We find \cannon\ temperatures to be in agreement to 60\,K
compared to a subsample of 1,702 sources with color-derived temperatures, and 
find \cannon\ metallicites to be in agreement to 0.08\,dex metallicity
compared to a subsample of 15 FGK+M or M+M binaries. 
Finally, our comparison between \cannon\ and APOGEE pipeline (ASPCAP) labels finds that 
ASPCAP is systematically biased towards reporting higher temperatures and lower metallicities. 
\end{abstract}
